---
title: "Mis-selling, Lapse and Planner Settlement: Panel Analysis with Residual Diagnostics"
subtitle: "Korean insurers, company-level, 2021--2025 (FSS + KNIA disclosures)"
date: "2026-10-05"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
    code-fold: show
    fig-width: 7
    fig-height: 4.5
execute:
  warning: false
  message: false
---

# Overview

This document reproduces the company-level panel study and adds a full **residual analysis** of the
preferred two-way fixed-effects (TWFE) models. The raw data are hard-coded below, so the file runs on its own
with no downloads.

* **Sources.** FSS insurance company disclosure, "contract management": planner settlement rate and
  13th/25th-month persistency, calendar years 2021--2025. KNIA consumer portal: item 04, the mis-selling
  rate (half-year all-channel rates, averaged to years); item 07, the long-term claim denial rate and the
  post-claim cancellation rate, summed counts.
* **Units.** `settle`, `p13`, `p25`, `deny_rate` and `post_claim_cxl_rate` are in %. `ms` is in % (0.01 = 1 basis point).
  `claims` is the number of long-term claims (KNIA item 07).
* **Coverage.** KNIA covers non-life insurers only. Mis-selling is therefore `NA` for life insurers, and the
  life sample is used only for settlement vs. lapse.

# Setup

```{r setup}
pkgs <- c("plm", "lmtest", "sandwich")
for (p in pkgs) if (!requireNamespace(p, quietly = TRUE)) install.packages(p, repos = "https://cloud.r-project.org")
library(plm); library(lmtest); library(sandwich)
set.seed(20261005)
pal <- c(blue = "#2a78d6", orange = "#eb6834", aqua = "#1baf7a", yellow = "#eda100", grey = "#52514e")
op <- par(no.readonly = TRUE)
```

# Raw data (hard-coded)

Values are as published. Zeros that mean "not applicable" are cleaned in the next section, not here.

```{r raw-data}
raw <- read.csv(text = "
sector,company,year,settle,p13,p25,ms,deny_rate,post_claim_cxl_rate,claims
life,ABL,2021,60.0,81.8,68.3,NA,NA,NA,NA
life,ABL,2022,64.2,81.6,65.5,NA,NA,NA,NA
life,ABL,2023,74.6,85.8,58.1,NA,NA,NA,NA
life,ABL,2024,81.5,89.6,69.0,NA,NA,NA,NA
life,ABL,2025,80.1,90.0,77.7,NA,NA,NA,NA
life,AIA,2021,30.5,79.7,63.9,NA,NA,NA,NA
life,AIA,2022,30.5,79.2,65.4,NA,NA,NA,NA
life,AIA,2023,19.6,80.3,64.5,NA,NA,NA,NA
life,AIA,2024,34.7,81.9,64.9,NA,NA,NA,NA
life,AIA,2025,30.7,81.3,66.3,NA,NA,NA,NA
life,BNP_Cardif_Life,2021,NA,68.6,31.5,NA,NA,NA,NA
life,BNP_Cardif_Life,2022,NA,84.0,50.6,NA,NA,NA,NA
life,BNP_Cardif_Life,2023,NA,77.1,59.3,NA,NA,NA,NA
life,BNP_Cardif_Life,2024,NA,84.5,49.1,NA,NA,NA,NA
life,BNP_Cardif_Life,2025,NA,68.5,69.4,NA,NA,NA,NA
life,DB_Life,2021,34.9,82.8,62.0,NA,NA,NA,NA
life,DB_Life,2022,40.4,85.4,63.0,NA,NA,NA,NA
life,DB_Life,2023,35.3,87.5,69.2,NA,NA,NA,NA
life,DB_Life,2024,45.2,90.3,72.0,NA,NA,NA,NA
life,DB_Life,2025,43.5,91.1,77.8,NA,NA,NA,NA
life,IBK_Pension,2021,NA,87.1,77.0,NA,NA,NA,NA
life,IBK_Pension,2022,NA,86.4,76.0,NA,NA,NA,NA
life,IBK_Pension,2023,NA,76.1,67.8,NA,NA,NA,NA
life,IBK_Pension,2024,NA,87.8,65.0,NA,NA,NA,NA
life,IBK_Pension,2025,NA,91.2,73.3,NA,NA,NA,NA
life,KB_Life,2021,20.8,80.4,63.1,NA,NA,NA,NA
life,KB_Life,2022,NA,89.6,69.1,NA,NA,NA,NA
life,KB_Life,2023,NA,78.1,55.3,NA,NA,NA,NA
life,KB_Life,2024,NA,89.9,61.2,NA,NA,NA,NA
life,KB_Life,2025,33.3,86.5,75.7,NA,NA,NA,NA
life,KDB_Life,2021,29.5,81.1,67.5,NA,NA,NA,NA
life,KDB_Life,2022,24.4,83.9,66.8,NA,NA,NA,NA
life,KDB_Life,2023,31.4,87.4,69.6,NA,NA,NA,NA
life,KDB_Life,2024,21.4,89.3,74.3,NA,NA,NA,NA
life,KDB_Life,2025,29.2,90.5,77.6,NA,NA,NA,NA
life,iM_Life,2021,7.1,80.5,65.8,NA,NA,NA,NA
life,iM_Life,2022,NA,86.2,67.7,NA,NA,NA,NA
life,iM_Life,2023,NA,88.2,77.5,NA,NA,NA,NA
life,iM_Life,2024,NA,91.6,67.4,NA,NA,NA,NA
life,iM_Life,2025,NA,91.8,81.4,NA,NA,NA,NA
life,Kyobo,2021,41.2,82.3,65.2,NA,NA,NA,NA
life,Kyobo,2022,39.3,81.7,66.4,NA,NA,NA,NA
life,Kyobo,2023,40.8,76.4,46.8,NA,NA,NA,NA
life,Kyobo,2024,47.3,88.2,64.7,NA,NA,NA,NA
life,Kyobo,2025,49.4,90.2,75.0,NA,NA,NA,NA
life,Kyobo_LifePlanet,2021,NA,88.2,80.1,NA,NA,NA,NA
life,Kyobo_LifePlanet,2022,NA,88.1,82.2,NA,NA,NA,NA
life,Kyobo_LifePlanet,2023,NA,65.0,28.4,NA,NA,NA,NA
life,Kyobo_LifePlanet,2024,NA,91.9,57.1,NA,NA,NA,NA
life,Kyobo_LifePlanet,2025,NA,85.0,88.2,NA,NA,NA,NA
life,NH_Life,2021,26.6,83.2,69.0,NA,NA,NA,NA
life,NH_Life,2022,18.8,83.0,71.9,NA,NA,NA,NA
life,NH_Life,2023,21.2,82.9,69.1,NA,NA,NA,NA
life,NH_Life,2024,37.2,88.8,74.6,NA,NA,NA,NA
life,NH_Life,2025,43.1,92.6,83.6,NA,NA,NA,NA
life,Tongyang,2021,29.4,84.2,61.7,NA,NA,NA,NA
life,Tongyang,2022,27.9,83.7,64.8,NA,NA,NA,NA
life,Tongyang,2023,32.0,87.8,59.2,NA,NA,NA,NA
life,Tongyang,2024,62.4,88.8,78.3,NA,NA,NA,NA
life,Tongyang,2025,57.0,89.9,78.6,NA,NA,NA,NA
life,Lina_Life,2021,38.4,85.5,77.0,NA,NA,NA,NA
life,Lina_Life,2022,31.1,85.6,77.7,NA,NA,NA,NA
life,Lina_Life,2023,3.2,83.9,74.3,NA,NA,NA,NA
life,Lina_Life,2024,17.0,82.3,70.8,NA,NA,NA,NA
life,Lina_Life,2025,23.7,81.8,70.4,NA,NA,NA,NA
life,MetLife,2021,25.1,80.5,62.7,NA,NA,NA,NA
life,MetLife,2022,24.5,84.8,65.7,NA,NA,NA,NA
life,MetLife,2023,21.8,85.9,68.8,NA,NA,NA,NA
life,MetLife,2024,32.2,88.1,68.6,NA,NA,NA,NA
life,MetLife,2025,48.5,88.4,76.0,NA,NA,NA,NA
life,MiraeAsset,2021,52.3,79.8,61.2,NA,NA,NA,NA
life,MiraeAsset,2022,NA,83.5,66.1,NA,NA,NA,NA
life,MiraeAsset,2023,NA,82.4,68.1,NA,NA,NA,NA
life,MiraeAsset,2024,NA,80.8,63.1,NA,NA,NA,NA
life,MiraeAsset,2025,NA,79.5,65.7,NA,NA,NA,NA
life,Samsung_Life,2021,49.8,88.3,68.8,NA,NA,NA,NA
life,Samsung_Life,2022,47.2,90.2,75.2,NA,NA,NA,NA
life,Samsung_Life,2023,43.1,85.4,65.5,NA,NA,NA,NA
life,Samsung_Life,2024,48.4,89.7,68.9,NA,NA,NA,NA
life,Samsung_Life,2025,49.2,88.0,77.2,NA,NA,NA,NA
life,Shinhan_Life,2021,30.9,81.4,62.5,NA,NA,NA,NA
life,Shinhan_Life,2022,34.7,82.0,64.6,NA,NA,NA,NA
life,Shinhan_Life,2023,31.7,84.1,63.0,NA,NA,NA,NA
life,Shinhan_Life,2024,44.8,87.0,69.4,NA,NA,NA,NA
life,Shinhan_Life,2025,36.9,88.1,75.6,NA,NA,NA,NA
life,Chubb_Life,2021,31.9,64.9,34.5,NA,NA,NA,NA
life,Chubb_Life,2022,18.1,68.9,45.2,NA,NA,NA,NA
life,Chubb_Life,2023,NA,81.9,62.0,NA,NA,NA,NA
life,Chubb_Life,2024,NA,89.4,68.8,NA,NA,NA,NA
life,Chubb_Life,2025,NA,91.7,80.1,NA,NA,NA,NA
life,Prudential,2021,46.3,84.0,72.4,NA,NA,NA,NA
life,Fubon_Hyundai,2021,32.8,81.5,68.7,NA,NA,NA,NA
life,Fubon_Hyundai,2022,21.0,85.8,68.5,NA,NA,NA,NA
life,Fubon_Hyundai,2023,28.6,78.1,64.6,NA,NA,NA,NA
life,Fubon_Hyundai,2024,28.2,87.6,60.3,NA,NA,NA,NA
life,Fubon_Hyundai,2025,27.0,87.4,77.7,NA,NA,NA,NA
life,Hana_Life,2021,11.8,91.2,72.8,NA,NA,NA,NA
life,Hana_Life,2022,5.0,79.4,75.1,NA,NA,NA,NA
life,Hana_Life,2023,48.3,82.4,62.7,NA,NA,NA,NA
life,Hana_Life,2024,10.8,90.3,65.0,NA,NA,NA,NA
life,Hana_Life,2025,NA,88.6,80.4,NA,NA,NA,NA
life,Hanwha_Life,2021,45.5,84.8,64.6,NA,NA,NA,NA
life,Hanwha_Life,2022,NA,84.2,68.6,NA,NA,NA,NA
life,Hanwha_Life,2023,NA,87.6,59.2,NA,NA,NA,NA
life,Hanwha_Life,2024,NA,88.0,72.8,NA,NA,NA,NA
life,Hanwha_Life,2025,NA,89.7,75.1,NA,NA,NA,NA
life,Heungkuk_Life,2021,38.5,85.8,69.8,NA,NA,NA,NA
life,Heungkuk_Life,2022,21.8,84.3,71.8,NA,NA,NA,NA
life,Heungkuk_Life,2023,22.4,76.2,45.5,NA,NA,NA,NA
life,Heungkuk_Life,2024,12.3,85.2,60.8,NA,NA,NA,NA
life,Heungkuk_Life,2025,29.6,84.9,72.8,NA,NA,NA,NA
nonlife,AIG,2020,NA,NA,NA,0.17,2.235,0.0711,80046
nonlife,AIG,2021,45.6,70.1,57.5,0.15,3.3529,0.0802,79632
nonlife,AIG,2022,25.2,73.4,56.9,0.125,2.9905,0.0429,79853
nonlife,AIG,2023,28.9,69.7,57.5,0.11,1.8177,0.0477,77239
nonlife,AIG,2024,12.4,62.9,53.9,0.1,2.0254,0.0,56284
nonlife,AIG,2025,43.8,60.8,46.0,0.08,1.6357,0.0422,45729
nonlife,AXA,2020,NA,NA,NA,0.015,1.576,0.3082,23477
nonlife,AXA,2021,0.0,76.5,66.2,0.025,1.7377,0.1925,29579
nonlife,AXA,2022,27.1,69.6,66.6,0.015,1.7038,0.1538,33043
nonlife,AXA,2023,10.7,75.6,57.5,0.02,1.6049,0.1911,36576
nonlife,AXA,2024,7.7,72.1,60.7,0.02,1.6474,0.1499,40305
nonlife,AXA,2025,15.4,70.9,56.4,0.015,1.7001,0.1192,43292
nonlife,DB_Ins,2020,NA,NA,NA,0.035,1.4852,0.1725,1055955
nonlife,DB_Ins,2021,70.4,88.7,74.2,0.03,1.5558,0.1539,1357218
nonlife,DB_Ins,2022,64.4,90.1,76.2,0.02,1.5657,0.1509,1622444
nonlife,DB_Ins,2023,62.7,88.9,75.8,0.02,1.6118,0.1862,1885259
nonlife,DB_Ins,2024,65.4,88.5,74.1,0.01,1.4185,0.2327,1847456
nonlife,DB_Ins,2025,64.4,88.8,73.6,0.01,1.3444,0.2017,2094304
nonlife,KB_Ins,2020,NA,NA,NA,0.025,1.0496,0.172,789159
nonlife,KB_Ins,2021,57.2,86.5,68.7,0.02,1.6338,0.1925,952056
nonlife,KB_Ins,2022,60.8,89.4,71.8,0.01,1.6188,0.1461,1087954
nonlife,KB_Ins,2023,55.6,87.6,74.6,0.01,1.4957,0.1392,1314126
nonlife,KB_Ins,2024,54.9,88.1,70.2,0.01,1.4932,0.2225,1245205
nonlife,KB_Ins,2025,66.7,88.5,71.9,0.01,1.3779,0.1907,1596248
nonlife,NH_Ins,2020,NA,NA,NA,0.03,0.7966,0.3176,113477
nonlife,NH_Ins,2021,50.0,91.3,75.0,0.01,0.5892,0.3598,120503
nonlife,NH_Ins,2022,42.5,85.9,85.5,0.01,0.4789,0.401,123187
nonlife,NH_Ins,2023,42.3,81.0,74.7,0.01,1.0797,0.5802,135316
nonlife,NH_Ins,2024,50.7,82.8,71.6,0.01,1.669,0.7853,117199
nonlife,NH_Ins,2025,46.0,85.6,71.6,0.01,1.6054,1.1456,116608
nonlife,Lotte,2020,NA,NA,NA,0.055,1.3346,0.1678,157795
nonlife,Lotte,2021,60.6,89.3,79.1,0.07,1.6115,0.1352,195531
nonlife,Lotte,2022,53.7,89.4,83.3,0.035,1.6089,0.1325,238485
nonlife,Lotte,2023,48.8,88.4,76.6,0.02,1.4318,0.1124,292157
nonlife,Lotte,2024,36.7,85.7,70.1,0.02,1.5575,0.1217,302278
nonlife,Lotte,2025,23.5,80.0,64.9,0.02,1.362,0.1002,283553
nonlife,Meritz,2020,NA,NA,NA,0.035,1.5328,0.1477,1419599
nonlife,Meritz,2021,45.3,83.6,69.4,0.025,1.6207,0.121,1455630
nonlife,Meritz,2022,45.8,84.6,69.2,0.01,1.5829,0.0709,1263618
nonlife,Meritz,2023,45.1,84.7,66.5,0.01,1.5202,0.1204,1238376
nonlife,Meritz,2024,48.0,84.4,64.6,NA,1.5491,0.1834,1045025
nonlife,Meritz,2025,40.4,84.4,64.5,0.01,1.3968,0.2039,1151132
nonlife,Samsung_FM,2020,NA,NA,NA,0.025,1.5822,0.1516,1182241
nonlife,Samsung_FM,2021,66.3,86.8,61.1,0.02,1.6603,0.1503,1474270
nonlife,Samsung_FM,2022,44.3,87.6,69.7,0.02,1.578,0.1618,1550960
nonlife,Samsung_FM,2023,47.8,86.3,68.5,0.02,1.4606,0.1818,1810869
nonlife,Samsung_FM,2024,65.9,88.2,67.2,0.01,1.3958,0.2276,1835860
nonlife,Samsung_FM,2025,75.1,89.5,72.4,0.01,1.3286,0.1585,2440415
nonlife,ShinhanEZ,2020,NA,NA,NA,0.09,1.2821,0.4808,468
nonlife,ShinhanEZ,2021,NA,NA,NA,NA,2.1944,0.3497,319
nonlife,ShinhanEZ,2022,0.0,0.0,64.2,NA,1.3423,0.0,149
nonlife,ShinhanEZ,2023,NA,NA,NA,0.06,0.0,0.0,33
nonlife,ShinhanEZ,2024,NA,68.1,NA,0.05,1.4925,1.676,268
nonlife,ShinhanEZ,2025,NA,79.7,49.7,0.0,2.9324,1.2441,3683
nonlife,ACE_Lina,2020,NA,NA,NA,0.325,1.6989,0.1398,125902
nonlife,ACE_Lina,2021,NA,64.4,55.3,0.245,1.1272,0.1255,120027
nonlife,ACE_Lina,2022,0.0,63.3,50.2,0.235,1.1101,0.1221,92060
nonlife,ACE_Lina,2023,NA,65.4,47.3,0.08,1.5766,0.1962,61396
nonlife,ACE_Lina,2024,NA,68.0,47.3,0.055,2.2059,0.2154,53312
nonlife,ACE_Lina,2025,NA,64.6,50.1,0.07,2.0021,0.2371,61986
nonlife,Yebyeol_MG,2020,NA,NA,NA,0.02,1.3191,0.3395,51399
nonlife,Yebyeol_MG,2021,70.0,87.2,73.3,0.025,1.2286,0.2002,58198
nonlife,Yebyeol_MG,2022,57.7,86.7,76.0,0.01,0.9074,0.3001,62269
nonlife,Yebyeol_MG,2023,75.0,84.3,71.5,0.01,0.7286,0.1467,74806
nonlife,Yebyeol_MG,2024,71.4,86.5,70.2,0.0,0.8476,0.1209,80459
nonlife,Yebyeol_MG,2025,20.0,79.4,62.5,0.005,0.5979,0.1126,90490
nonlife,Cardif_GI,2021,NA,73.8,57.4,NA,NA,NA,NA
nonlife,KakaoPay,2024,NA,NA,NA,0.0,2.7053,0.0396,3105
nonlife,KakaoPay,2025,NA,NA,NA,0.005,1.7519,0.0851,37045
nonlife,Hana_Ins,2020,NA,NA,NA,0.03,1.703,0.5576,5226
nonlife,Hana_Ins,2021,47.0,88.0,78.2,0.02,2.8995,1.0068,5863
nonlife,Hana_Ins,2022,28.1,87.2,79.4,0.025,2.5589,1.6575,16843
nonlife,Hana_Ins,2023,37.8,87.6,75.2,0.01,2.155,1.192,34107
nonlife,Hana_Ins,2024,47.2,83.3,76.7,0.025,1.8735,1.142,51455
nonlife,Hana_Ins,2025,26.6,79.6,69.7,0.015,1.8797,1.0805,58200
nonlife,Hanwha_GI,2020,NA,NA,NA,0.06,1.5143,0.1538,485032
nonlife,Hanwha_GI,2021,55.7,81.1,66.4,0.035,1.5331,0.1796,397548
nonlife,Hanwha_GI,2022,63.6,84.3,63.6,0.03,1.3308,0.1758,362330
nonlife,Hanwha_GI,2023,66.2,84.9,66.4,0.025,0.7669,0.1022,482615
nonlife,Hanwha_GI,2024,74.9,86.4,67.5,0.03,0.9475,0.1135,564964
nonlife,Hanwha_GI,2025,74.1,85.1,71.1,0.02,0.9737,0.1386,765849
nonlife,Hyundai_MF,2020,NA,NA,NA,0.035,1.9594,0.1577,1386058
nonlife,Hyundai_MF,2021,66.6,87.4,73.0,0.03,1.7859,0.1675,1794348
nonlife,Hyundai_MF,2022,57.6,88.8,74.9,0.02,1.5044,0.1702,2136072
nonlife,Hyundai_MF,2023,58.0,87.0,75.6,0.02,1.4245,0.2023,2567540
nonlife,Hyundai_MF,2024,57.0,87.1,73.3,0.015,1.5926,0.1774,2300898
nonlife,Hyundai_MF,2025,54.1,86.8,73.4,0.01,1.3303,0.174,2412342
nonlife,Heungkuk_FM,2020,NA,NA,NA,0.065,1.3844,0.1573,106615
nonlife,Heungkuk_FM,2021,63.3,86.0,70.7,0.04,1.5666,0.1498,141454
nonlife,Heungkuk_FM,2022,55.7,88.6,74.5,0.025,1.6317,0.1228,199429
nonlife,Heungkuk_FM,2023,60.6,88.0,76.6,0.02,1.2976,0.1314,297478
nonlife,Heungkuk_FM,2024,75.0,86.6,73.5,0.01,1.2178,0.1098,376759
nonlife,Heungkuk_FM,2025,68.2,86.0,71.1,0.015,1.1147,0.1368,466306
", stringsAsFactors = FALSE)
str(raw)
table(raw$sector, raw$year)
```

# Cleaning and variable construction

* A published value of exactly 0.0 for settlement or persistency means the company had no newly registered
  exclusive planners, or no reportable cohort, in the window. These are set to `NA`.
* Lapse $= 100 -$ persistency. Mis-selling is rescaled to basis points.
* **Timing.** Persistency published for year $t$ refers to contracts written in $t-1$ (13th month) or $t-2$
  (25th month). We therefore use $\text{MS}_{t-1}$ for 13th-month lapse, and $\text{MS}_{t-2}$ and
  $\text{Settle}_{t-1}$ for 25th-month lapse.
* Lags are true one-year lags: the panel is completed to a full company x year grid before lagging.

```{r clean}
d <- raw
for (v in c("settle", "p13", "p25")) d[[v]][!is.na(d[[v]]) & d[[v]] == 0] <- NA
d$lapse13 <- 100 - d$p13
d$lapse25 <- 100 - d$p25
d$ms_bp   <- 100 * d$ms

# complete company x year grid, then lag within company
pan <- do.call(rbind, lapply(split(d, d$company), function(g) {
  yrs <- seq(min(g$year), max(g$year))
  out <- merge(data.frame(company = g$company[1], sector = g$sector[1], year = yrs),
               g[, setdiff(names(g), c("company", "sector"))], by = "year", all.x = TRUE)
  out <- out[order(out$year), ]
  lag1 <- function(x) c(NA, head(x, -1))
  out$ms_bp_l1  <- lag1(out$ms_bp)
  out$ms_bp_l2  <- lag1(out$ms_bp_l1)
  out$settle_l1 <- lag1(out$settle)
  out
}))
rownames(pan) <- NULL

NONFACE <- c("AIG", "ACE_Lina", "AXA")             # telemarketing / direct-led insurers
EXCL    <- c("ShinhanEZ", "KakaoPay", "Cardif_GI") # digital or too sparsely reported

nl <- subset(pan, sector == "nonlife" & year >= 2021 & year <= 2025 & !company %in% EXCL)
lf <- subset(pan, sector == "life"    & year >= 2021 & year <= 2025)
nl$nonface <- nl$company %in% NONFACE
c(nonlife_firms = length(unique(nl$company)), nonlife_obs = nrow(nl),
  life_firms = length(unique(lf$company)), life_obs = nrow(lf))
```

# Descriptive statistics

The between SD is the SD of firm means; the within SD is the SD of deviations from firm means.

```{r describe}
describe <- function(df, vars) {
  do.call(rbind, lapply(vars, function(v) {
    x <- df[!is.na(df[[v]]), c("company", v)]
    m <- ave(x[[v]], x$company)
    data.frame(var = v, N = nrow(x), firms = length(unique(x$company)),
               mean = mean(x[[v]]), sd = sd(x[[v]]),
               sd_between = sd(tapply(x[[v]], x$company, mean)),
               sd_within = sd(x[[v]] - m))
  }))
}
knitr_ok <- requireNamespace("knitr", quietly = TRUE)
show <- function(x, digits = 3) if (knitr_ok) knitr::kable(x, digits = digits) else print(x, digits = digits)
show(describe(nl, c("lapse13", "lapse25", "settle", "ms_bp", "deny_rate", "post_claim_cxl_rate")), 2)
show(describe(lf, c("lapse13", "lapse25", "settle")), 2)
```

```{r ms-trend, fig.cap="Mis-selling rate by non-life insurer. TM/direct-led insurers in orange."}
ms_all <- subset(pan, sector == "nonlife" & !company %in% EXCL)
plot(NA, xlim = c(2020, 2025), ylim = c(0, max(ms_all$ms_bp, na.rm = TRUE)),
     xlab = "Year", ylab = "Mis-selling rate (bp)", las = 1)
grid(col = "#e1e0d9")
for (co in unique(ms_all$company)) {
  g <- ms_all[ms_all$company == co, ]
  lines(g$year, g$ms_bp, col = if (co %in% NONFACE) pal["orange"] else pal["blue"], lwd = 1.4)
}
legend("topright", c("TM/direct-led", "Face-to-face-led"), col = pal[c("orange", "blue")], lwd = 2, bty = "n")
```

# Correlation decomposition

* **Pooled** correlation: all firm-years together.
* **Between** correlation: computed on firm means.
* **Within** correlation: computed after removing firm and year means (two-way demeaning).

```{r corr}
twoway_demean <- function(x, id, t) x - ave(x, id) - ave(x, t) + mean(x)
corr_decomp <- function(df, x, y) {
  z <- df[complete.cases(df[, c(x, y)]), c("company", "year", x, y)]
  bm <- aggregate(z[, c(x, y)], list(company = z$company), mean)
  wx <- twoway_demean(z[[x]], z$company, z$year); wy <- twoway_demean(z[[y]], z$company, z$year)
  data.frame(x = x, y = y, N = nrow(z), firms = nrow(bm),
             r_pooled = cor(z[[x]], z[[y]]), p_pooled = cor.test(z[[x]], z[[y]])$p.value,
             r_between = cor(bm[[x]], bm[[y]]), p_between = cor.test(bm[[x]], bm[[y]])$p.value,
             r_within = cor(wx, wy), p_within = cor.test(wx, wy)$p.value)
}
cd <- rbind(
  cbind(sample = "nonlife", corr_decomp(nl, "ms_bp_l1", "lapse13")),
  cbind(sample = "nonlife", corr_decomp(nl, "ms_bp_l2", "lapse25")),
  cbind(sample = "nonlife", corr_decomp(nl, "settle", "lapse13")),
  cbind(sample = "nonlife", corr_decomp(nl, "settle_l1", "lapse25")),
  cbind(sample = "nonlife", corr_decomp(nl, "settle", "ms_bp")),
  cbind(sample = "nonlife", corr_decomp(nl, "deny_rate", "lapse13")),
  cbind(sample = "life",    corr_decomp(lf, "settle", "lapse13")),
  cbind(sample = "life",    corr_decomp(lf, "settle_l1", "lapse25")))
show(cd, 3)
```

The within-correlation p-values ignore the degrees of freedom absorbed by demeaning, so treat them as indicative only.

# Panel regressions

The specifications are:

* **A** (non-life): $\text{Lapse13}_{it} = \alpha_i + \lambda_t + \beta\,\text{MS}_{i,t-1} + \gamma\,\text{Settle}_{it} + \varepsilon_{it}$
* **B** (non-life): $\text{Lapse25}_{it} = \alpha_i + \lambda_t + \beta\,\text{MS}_{i,t-2} + \gamma\,\text{Settle}_{i,t-1} + \varepsilon_{it}$
* **C** (non-life): $\text{MS}_{it} = \alpha_i + \lambda_t + \gamma\,\text{Settle}_{it} + \varepsilon_{it}$
* **L1** (life): $\text{Lapse13}_{it} = \alpha_i + \lambda_t + \gamma\,\text{Settle}_{it} + \varepsilon_{it}$

Each specification is estimated as pooled OLS, random effects, firm FE and two-way FE.
Standard errors are firm-clustered (Arellano HC1).

```{r models}
pnl <- pdata.frame(nl, index = c("company", "year"))
plf <- pdata.frame(lf, index = c("company", "year"))

fit_all <- function(f, pd) list(
  pooled = plm(f, pd, model = "pooling"),
  random = plm(f, pd, model = "random"),
  fe     = plm(f, pd, model = "within", effect = "individual"),
  twfe   = plm(f, pd, model = "within", effect = "twoways"))
cl <- function(m) coeftest(m, vcov = vcovHC(m, method = "arellano", type = "HC1", cluster = "group"))

fA <- lapse13 ~ ms_bp_l1 + settle
fB <- lapse25 ~ ms_bp_l2 + settle_l1
fC <- ms_bp ~ settle
fL <- lapse13 ~ settle
mA <- fit_all(fA, pnl); mB <- fit_all(fB, pnl); mC <- fit_all(fC, pnl); mL <- fit_all(fL, plf)

tab <- function(ms, vars) {
  do.call(rbind, lapply(names(ms), function(k) {
    ct <- cl(ms[[k]])
    do.call(rbind, lapply(vars, function(v) if (v %in% rownames(ct))
      data.frame(model = k, term = v, estimate = ct[v, 1], cluster_se = ct[v, 2], p = ct[v, 4],
                 N = nobs(ms[[k]]))))
  }))
}
show(tab(mA, c("ms_bp_l1", "settle")), 3)
show(tab(mB, c("ms_bp_l2", "settle_l1")), 3)
show(tab(mC, "settle"), 3)
show(tab(mL, "settle"), 3)
```

## Hausman tests (FE vs. RE)

```{r hausman}
rbind(A  = unlist(phtest(mA$fe, mA$random)[c("statistic", "p.value")]),
      B  = unlist(phtest(mB$fe, mB$random)[c("statistic", "p.value")]),
      C  = unlist(phtest(mC$fe, mC$random)[c("statistic", "p.value")]),
      L1 = unlist(phtest(mL$fe, mL$random)[c("statistic", "p.value")]))
```

A rejection means the firm effects are correlated with the regressors, so RE and pooled estimates are biased and FE is preferred.

## Wild cluster bootstrap (few clusters)

The non-life sample has only about 13 firm clusters. Cluster-robust t-tests over-reject in that setting, so the
key two-way FE coefficients are re-tested with a wild cluster *restricted* bootstrap. It uses Webb six-point
weights, imposes the null, and is estimated as least-squares dummy variables (LSDV).

```{r wcr}
wcr_boot <- function(df, y, xs, test, B = 9999) {
  z <- df[complete.cases(df[, c(y, xs)]), ]
  X <- model.matrix(reformulate(c(xs, "factor(company)", "factor(year)")), z)
  Y <- z[[y]]; g <- as.integer(factor(z$company)); G <- max(g)
  k <- which(colnames(X) == test)
  tstat <- function(Yv) {
    XtXi <- MASS_ginv(crossprod(X)); b <- XtXi %*% crossprod(X, Yv); u <- as.vector(Yv - X %*% b)
    meat <- Reduce(`+`, lapply(seq_len(G), function(j) { s <- crossprod(X[g == j, , drop = FALSE], u[g == j]); s %*% t(s) }))
    n <- nrow(X); p <- qr(X)$rank
    V <- (G / (G - 1)) * ((n - 1) / (n - p)) * XtXi %*% meat %*% XtXi
    c(t = b[k] / sqrt(V[k, k]), b = b[k])
  }
  t0 <- tstat(Y)
  Xr <- X[, -k, drop = FALSE]; br <- qr.coef(qr(Xr), Y); br[is.na(br)] <- 0
  fitted_r <- as.vector(Xr %*% br); ur <- Y - fitted_r
  webb <- c(-sqrt(1.5), -1, -sqrt(0.5), sqrt(0.5), 1, sqrt(1.5))
  tb <- replicate(B, tstat(fitted_r + sample(webb, G, replace = TRUE)[g] * ur)["t"])
  c(estimate = unname(t0["b"]), t = unname(t0["t"]), p_boot = mean(abs(tb) >= abs(t0["t"])), clusters = G)
}
MASS_ginv <- function(A) { s <- svd(A); tol <- max(dim(A)) * max(s$d) * .Machine$double.eps
  s$v %*% (ifelse(s$d > tol, 1 / s$d, 0) * t(s$u)) }

B_DRAWS <- 1999   # raise to 9999 for publication-quality p-values
rbind(
  A_ms_bp_l1   = wcr_boot(nl, "lapse13", c("ms_bp_l1", "settle"), "ms_bp_l1", B_DRAWS),
  A_settle     = wcr_boot(nl, "lapse13", c("ms_bp_l1", "settle"), "settle", B_DRAWS),
  B_settle_l1  = wcr_boot(nl, "lapse25", c("ms_bp_l2", "settle_l1"), "settle_l1", B_DRAWS),
  C_settle     = wcr_boot(nl, "ms_bp", "settle", "settle", B_DRAWS),
  L1_settle    = wcr_boot(lf, "lapse13", "settle", "settle", B_DRAWS))
```

# Residual analysis

Residual diagnostics are run on the **LSDV** equivalent of each two-way FE model: `lm` with firm and year dummies.
It gives identical slope coefficients to `plm`'s within estimator and gives access to fitted values, leverage,
studentised residuals and Cook's distance. The first check confirms the two estimators agree.

```{r lsdv}
lsdv <- function(f, df) {
  z <- df[complete.cases(df[, all.vars(f)]), ]
  z <- z[ave(z$year, z$company, FUN = length) >= 2, ]   # singleton firms have leverage 1 and carry no FE information
  m <- lm(update(f, . ~ . + factor(company) + factor(year)), data = z)
  m$panel <- z; m
}
lA <- lsdv(fA, nl); lB <- lsdv(fB, nl); lC <- lsdv(fC, nl); lL <- lsdv(fL, lf)
rbind(plm_twfe = coef(mA$twfe), lm_lsdv = coef(lA)[names(coef(mA$twfe))])
rbind(plm_twfe_L1 = coef(mL$twfe), lm_lsdv_L1 = coef(lL)["settle"])   # identical after dropping singletons
```

## Why fixed effects? Pooled residuals cluster by firm

If the pooled model were adequate, each firm's residuals would centre on zero. Instead they shift by firm,
which shows unobserved firm heterogeneity. The firm FE models absorb it.

```{r pooled-vs-fe, fig.height=5}
zA <- lA$panel
pooledA <- lm(fA, data = zA)
par(mfrow = c(1, 2), mar = c(7, 4, 2, 1))
boxplot(resid(pooledA) ~ zA$company, las = 2, xlab = "", ylab = "Residual (pp)", main = "Pooled OLS",
        col = ifelse(sort(unique(zA$company)) %in% NONFACE, pal["orange"], pal["blue"]), cex.axis = 0.7)
abline(h = 0, lty = 2)
boxplot(resid(lA) ~ zA$company, las = 2, xlab = "", ylab = "Residual (pp)", main = "Two-way FE",
        col = ifelse(sort(unique(zA$company)) %in% NONFACE, pal["orange"], pal["blue"]), cex.axis = 0.7)
abline(h = 0, lty = 2)
par(op)
# share of residual variance explained by firm identity
c(pooled = summary(lm(resid(pooledA) ~ zA$company))$r.squared,
  twfe   = summary(lm(resid(lA) ~ zA$company))$r.squared)
```

## Standard diagnostic plots (spec A, 13th-month lapse)

```{r diag-plots, fig.height=7}
diag_panel <- function(m, title) {
  z <- m$panel; r <- resid(m); f <- fitted(m); rs <- rstudent(m)
  col <- ifelse(z$company %in% NONFACE, pal["orange"], pal["blue"])
  par(mfrow = c(2, 2), mar = c(4, 4, 2.5, 1), oma = c(0, 0, 1.5, 0))
  plot(f, r, pch = 19, col = col, xlab = "Fitted", ylab = "Residual", main = "Residuals vs fitted")
  abline(h = 0, lty = 2); lines(lowess(f, r), col = pal["grey"], lwd = 2)
  qqnorm(rs, pch = 19, col = col, main = "Normal Q-Q (studentised)"); qqline(rs, lty = 2)
  plot(f, sqrt(abs(rs)), pch = 19, col = col, xlab = "Fitted", ylab = expression(sqrt("|studentised residual|")),
       main = "Scale-location"); lines(lowess(f, sqrt(abs(rs))), col = pal["grey"], lwd = 2)
  h <- hatvalues(m); cdist <- cooks.distance(m)
  plot(h, rs, pch = 19, col = col, cex = 0.6 + 6 * cdist / max(cdist), xlab = "Leverage", ylab = "Studentised residual",
       main = "Influence (size = Cook's D)")
  abline(h = c(-2, 2), lty = 3)
  top <- order(cdist, decreasing = TRUE)[1:3]
  text(h[top], rs[top], paste(z$company[top], z$year[top]), pos = 4, cex = 0.7)
  mtext(title, outer = TRUE, font = 2)
  par(op)
}
diag_panel(lA, "Spec A: Lapse13 ~ MS(t-1) + Settle, two-way FE")
```

## Residuals by year and against regressors

```{r resid-structure, fig.height=6}
z <- lA$panel; r <- resid(lA)
par(mfrow = c(2, 2), mar = c(4, 4, 2.5, 1))
boxplot(r ~ z$year, xlab = "Year", ylab = "Residual", main = "By year", col = pal["aqua"]); abline(h = 0, lty = 2)
plot(z$settle, r, pch = 19, col = pal["blue"], xlab = "Settlement (%)", ylab = "Residual", main = "vs settlement")
abline(h = 0, lty = 2); lines(lowess(z$settle, r), col = pal["grey"], lwd = 2)
plot(z$ms_bp_l1, r, pch = 19, col = ifelse(z$company %in% NONFACE, pal["orange"], pal["blue"]),
     xlab = "Mis-selling t-1 (bp)", ylab = "Residual", main = "vs mis-selling t-1")
abline(h = 0, lty = 2)
# residual autocorrelation within firm: e_it vs e_i,t-1
z$r <- r; z <- z[order(z$company, z$year), ]
z$r_l1 <- ave(z$r, z$company, FUN = function(x) c(NA, head(x, -1)))
z$r_l1[c(FALSE, diff(z$year) != 1 | head(z$company, -1) != tail(z$company, -1))] <- NA
plot(z$r_l1, z$r, pch = 19, col = pal["blue"], xlab = "Residual t-1", ylab = "Residual t", main = "Serial dependence")
abline(h = 0, v = 0, lty = 2); abline(lm(r ~ r_l1, z), col = pal["grey"], lwd = 2)
par(op)
```

Within-firm demeaning induces a mechanical negative correlation between residuals in a short panel: about
$-1/(T-1) = -0.25$ for $T=5$. A lag coefficient near $-0.25$ is therefore the benchmark for "no serial
correlation", not zero. The Wooldridge test below uses that benchmark.

## Formal tests

```{r tests}
normality <- function(m) {
  r <- resid(m); n <- length(r); s <- (r - mean(r)) / sd(r)
  jb <- n / 6 * (mean(s^3)^2 + (mean(s^4) - 3)^2 / 4)
  c(shapiro_W = unname(shapiro.test(r)$statistic), shapiro_p = shapiro.test(r)$p.value,
    skew = mean(s^3), excess_kurt = mean(s^4) - 3, jarque_bera = jb, jb_p = 1 - pchisq(jb, 2))
}
hetero <- function(m, regs) {
  z <- m$panel
  bp_full <- bptest(m)                                         # all regressors incl. dummies
  bp_x    <- bptest(m, reformulate(regs), data = z)            # slopes only
  c(BP_all = unname(bp_full$statistic), BP_all_p = bp_full$p.value,
    BP_slopes = unname(bp_x$statistic), BP_slopes_p = bp_x$p.value)
}
spec_tests <- function(m) {
  rs <- resettest(m, power = 2:3, type = "fitted")
  c(RESET = unname(rs$statistic), RESET_p = rs$p.value)
}
panel_tests <- function(f, pd) {
  w  <- tryCatch(pwartest(f, data = pd), error = function(e) NULL)       # Wooldridge FE serial corr.
  cdt <- tryCatch(pcdtest(plm(f, pd, model = "within", effect = "twoways"), test = "cd"), error = function(e) NULL)
  c(Wooldridge_AR_F = if (!is.null(w)) unname(w$statistic) else NA,
    Wooldridge_AR_p = if (!is.null(w)) w$p.value else NA,
    Pesaran_CD = if (!is.null(cdt)) unname(cdt$statistic) else NA,
    Pesaran_CD_p = if (!is.null(cdt)) cdt$p.value else NA)
}
diag_tab <- rbind(
  A  = c(normality(lA), hetero(lA, c("ms_bp_l1", "settle")),   spec_tests(lA), panel_tests(fA, pnl)),
  B  = c(normality(lB), hetero(lB, c("ms_bp_l2", "settle_l1")), spec_tests(lB), panel_tests(fB, pnl)),
  C  = c(normality(lC), hetero(lC, "settle"),                   spec_tests(lC), panel_tests(fC, pnl)),
  L1 = c(normality(lL), hetero(lL, "settle"),                   spec_tests(lL), panel_tests(fL, plf)))
show(t(diag_tab), 3)
```

The tests are:

* **Shapiro--Wilk / Jarque--Bera**: normality of residuals. Inference does not rely on normality (clustered
  SE, bootstrap), but heavy tails flag outliers.
* **Breusch--Pagan**: heteroskedasticity, against all regressors and against the slope regressors only.
  Clustered SEs are robust to it either way.
* **RESET** (fitted$^2$, fitted$^3$): functional-form misspecification.
* **Wooldridge test**: serial correlation in FE residuals, accounting for the $-1/(T-1)$ demeaning bias.
* **Pesaran CD**: cross-sectional dependence, i.e. common shocks not absorbed. Year effects should remove most of it.

## Influence: leave-one-firm-out

With 13 clusters, a single insurer can drive a result. Below, each firm is dropped in turn and the TWFE model is re-estimated.

```{r jackknife, fig.height=4.5}
firms <- sort(unique(lA$panel$company))
jk <- t(sapply(firms, function(co) {
  m <- plm(fA, pdata.frame(subset(nl, company != co), index = c("company", "year")), model = "within", effect = "twoways")
  coef(m)[c("ms_bp_l1", "settle")]
}))
full <- coef(mA$twfe)[c("ms_bp_l1", "settle")]
show(cbind(jk, d_settle = jk[, "settle"] - full["settle"]), 3)
par(mfrow = c(1, 2), mar = c(7, 4, 2.5, 1))
for (v in c("settle", "ms_bp_l1")) {
  barplot(jk[, v], las = 2, cex.names = 0.7, col = ifelse(firms %in% NONFACE, pal["orange"], pal["blue"]),
          main = paste("Drop one firm:", v), ylab = "TWFE coefficient")
  abline(h = full[v], lty = 2, lwd = 2)
}
par(op)
range(jk[, "settle"])
```

## Influence: Cook's distance and outlier-trimmed refit

```{r cooks}
z <- lA$panel; cdist <- cooks.distance(lA); rs <- rstudent(lA)
flag <- cdist > 4 / nrow(z) | abs(rs) > 2.5
show(data.frame(company = z$company, year = z$year, lapse13 = z$lapse13, settle = z$settle,
                ms_bp_l1 = z$ms_bp_l1, rstudent = rs, cooks = cdist)[flag, ], 3)
trim <- z[!flag, ]
m_trim <- plm(fA, pdata.frame(trim, index = c("company", "year")), model = "within", effect = "twoways")
rbind(full = cl(mA$twfe)["settle", ], trimmed = cl(m_trim)["settle", ])
```

## Residuals of the life model (L1)

```{r life-diag, fig.height=7}
diag_panel(lL, "Spec L1 (life): Lapse13 ~ Settle, two-way FE")
```

# Summary

```{r summary}
ctA <- cl(mA$twfe); ctL <- cl(mL$twfe)
cat(sprintf(paste0(
  "Spec A (non-life, TWFE): settlement coef = %.3f (cluster SE %.3f, p = %.3f); ",
  "mis-selling(t-1) coef = %.3f (p = %.3f).\n",
  "Leave-one-firm-out range for settlement: [%.3f, %.3f].\n",
  "Spec L1 (life, TWFE): settlement coef = %.3f (p = %.3f).\n"),
  ctA["settle", 1], ctA["settle", 2], ctA["settle", 4], ctA["ms_bp_l1", 1], ctA["ms_bp_l1", 4],
  min(jk[, "settle"]), max(jk[, "settle"]), ctL["settle", 1], ctL["settle", 4]))
```

How to read the diagnostics:

1. **Firm heterogeneity is the first-order problem.** Pooled residuals are strongly firm-specific, and
   Hausman rejects RE for non-life. Two-way FE is the appropriate model.
2. **Residual distribution.** Check the Q-Q plot and the JB test for heavy tails. Any firm-years flagged in the
   Cook's distance table are re-tested in the trimmed refit.
3. **Heteroskedasticity, serial correlation and cross-sectional dependence.** Whatever the tests show, inference
   already uses firm-clustered SEs (robust to heteroskedasticity and within-firm serial correlation) and a wild
   cluster bootstrap (robust to few clusters). Year effects absorb common shocks.
4. **Robustness of the headline result.** The settlement effect on 13th-month lapse should be judged by the
   jackknife range and the trimmed refit. The mis-selling coefficient is not robust in any specification.


## Diagnostic findings for this data

These are the results produced by the hard-coded data above. The bootstrap p-values vary slightly with the number of draws.

| Check | Spec A: Lapse13 (non-life) | Spec C: Mis-selling | Spec L1: Lapse13 (life) |
|---|---|---|---|
| Firm heterogeneity | Firm identity explains 41% of pooled residual variance and 0% of TWFE residual variance; Hausman $\chi^2=80.5$, $p<0.001$ | Hausman $p=0.07$ | Hausman $p=0.47$ |
| Normality | Shapiro $p=0.10$, JB $p=0.46$: acceptable | Heavy tails (excess kurtosis 3.2, JB $p<0.001$) | Acceptable (JB $p=0.92$) |
| Heteroskedasticity (BP, slopes) | $p=0.03$ | $p=0.61$ | $p=0.03$ |
| Functional form (RESET) | $p=0.34$: fine | $p<0.001$: linear model inadequate | $p=0.26$ |
| Serial correlation (Wooldridge) | $p=0.03$ | $p<0.001$ | $p=0.16$ |
| Cross-sectional dependence (CD, two-way) | $p=0.63$: absorbed by year effects | $p=0.88$ | $p=0.36$ |
| Influence | Leave-one-firm-out settlement range $[-0.116,-0.076]$; never changes sign. Dropping 8 flagged firm-years gives $-0.123$ | -- | -- |

What this means:

* **Spec A (settlement and 13th-month lapse) is well specified.** The residuals are close to normal, the
  functional form is adequate and no single insurer drives the result. The heteroskedasticity and serial
  correlation found are handled by firm-clustered standard errors; the wild bootstrap gives $p \approx 0.05$.
  The settlement effect is negative in every leave-one-out fit and gets *stronger* when outliers are trimmed.
* **Spec C (mis-selling as outcome) fails the diagnostics.** It has heavy tails, fails RESET and shows strong
  serial correlation. Mis-selling is bounded at zero, rounded to 1 bp and mostly 1--2 bp, so a linear FE model
  is a poor description. Its null result reflects data limitations rather than a precise zero. A count or
  fractional-response model on the underlying cancellation counts would be the right next step if counts can be obtained.
* **Mis-selling in Spec A** is fragile. Dropping AIG alone moves its coefficient from $-0.70$ to $-0.06$.
* **Standard errors.** `plm`'s Arellano HC1 cluster SEs are slightly smaller than `linearmodels` in the PDF
  report (0.031 vs. 0.037 for settlement) because the two packages use different small-sample corrections.
  The bootstrap p-values agree.

```{r session}
sessionInfo()
```
